Unveiling Exception Handling Bug Hazards in Android based on GitHub and Google Code Issues

### 2015-03-27 android-stacks.tex
last changes

Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on its repository. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces.

Regardless of the issue labels, every exception stack
trace may contain relevant information concerning the exception structure of the
projects analyzed, and therefore can reveal \emph{bug hazards} on the exception-related code.
Because of this, we opted for not restricting the analysis to just defect issues.

@@

\subsection{Android Apps in Google Code}
Google Code contains widely used open-source Android apps (e.g. K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on the Google Code issue tracker: \url{https://github.com/k9mail/k-9/wiki/LoggingErrors}.}).
However, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects.\footnote{Google code used to provide a Web service to its repositories, but this was deactivated in June 2013 in what Google called a ``clean-up action''.}
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigates 
through the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.

To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
(on the Google Code search interface) for the term ``android''. On January 2014, when we queried Google Code, this resulted in a list of 788  projects. This list comprised the seeds sent to our Crawler.

The Crawler retrieved all issues and comments for these projects.
From this set, 724 projects defined at least 1 issue. Running the ExceptionMiner tool 
 in this set we observed that 183 projects had at least one issue containing an exception stack trace.
 Then we performed further clean up (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained  127,456 issues in total,
 from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the occurrences of different labels 
on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
the issues were labeled as ``Defect''. However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stack traces (regardless its labels).

@@

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace needs to be classified according to its type
(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of the exceptions reported 
on stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
a base exception type. Hence in this study the bytecode analysis was used to discover the type 
of each mined exception when the jar file of such exception was available on the project or 
on a reused  Java library. An specific implementation (based on source code analysis)
 was needed to discover the exception type when the bytecode was not available.

With this module we analyzed all exceptions defined in the Android platform (Version 4.4, API level 19),
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover, we also analyzed the 
exceptions reported on stack traces that were defined on applications and third-party libraries 
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e., Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristics adopted in this classification.

To conduct this classification, we provide this module with the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application.

To discover the packages for the first two origins
we can use to the Android specification.
To discover the packages for the third origin, the application itself, this module
extracts the manifest files of each Android app, which defines the main packages that the applications consist of.
 If this file is not available, the tool recursively analyzes the 
structure of source-code directories composing the appm filtering out the cases in which the application also includes the source code of reused libraries.
Then, based on this information and using pattern matching 
between the signaler name and the packages, this module identifies 
the origin of the exception signalers.

@@

 The NullPointerException was mainly signaled inside the application code (50\%) and the Android platform (31.5\%),
 although we could also find the NullPointerException being signaled by third-party libraries (16.3\%). 
Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or a bad practice to 
explicitly throw a NullPointerException. Some prefer to encapsulate such an exception on
an instance of IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
can be signalled by an API expecting a non-null argument.

The high prevalence of NullPointerException is aligned with the
findings of other 
research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher,kechagia2014}. 
For instance, Sunghun et
al.~\cite{kim2013predicting} show that 38\% the bugs related to
exception handling in the Eclipse project
 are caused by NullPointerException.
Furhtermore, Kechagia and Spinellis showed that the NullPointerException was the  
most reported exception on the crash reports sent to BugSense (a bug report 
management service for Android applications)  \cite{kechagia2014}.
Other research on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} shows that most of the automatically 
detected bugs were due to NullPointerException and exceptions
implicitly-signaled by the Java 
environment due to programming mistakes or resource limitations
 (as the ones found in our study).

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed a manual inspection in order to identify the underlying
concerns related to the most frequently reported root exceptions.
Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery)~\cite{Zhang12}.

Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
the concern identification of each exception was based on intended
meaning of the particular exception type, as defined in 
its Javadoc documentaion and in the Java specification. 
For example: (i) an instance of ArrayOutOfBoundException  
refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all 
exceptions related to backward compatibility~\cite{javaback}, such as
InstantiationError, VerifyError, and IllegalAccessError.

@@

This analysis revealed that the exceptions implicitly thrown by runtime environment 
due to mistakes on programming logic (e.g., out-of-bounds array index, null pointer 
references) and resource limitations represent a major bug hazard to the exception 
handling code of Android applications - security ..... os outros nao estao causando 
muitos problemas.

Although such exceptions do not violate the best practices described before (which are related to the 
explicitly thrown exceptions), they represent a major threat to apps robustness.

\finding{%
More than 50\% of the uncaught exceptions analyzed in this study are due to errors in
programming logic, with the NullPointerException as most prevalent
exception.
For another 25\% the root cause relates to resource constraints.
}

\subsection{Exception Types}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

We can observe that most of the reported exceptions are of type runtime 
(64.85\%); and that the most common origins are methods defined either on the Application (47.3\%)
or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).

We can also see, from Table~\ref{tab:typeroottab}, that in contrast with the other origins, most of the 
 exceptions signaled on Android Libcore (i.e., the set of libraries reused by Android) are 
checked exceptions. This set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
org.apache.http, org.json, org.xml, and javax. Signaling checked exceptions is considered a 
good practice (see best practice IV in Section~\ref{sec:best}) because by using 
checked exceptions a library can define a precise 
\emph{exception interface}~\cite{miller1997issues} to its clients.
 Since such libraries are widely used  in several projects, this
 finding can be attributed to the libraries' maturity.

@@

\finding{Almost two thirds of all crashes come from run time
  exceptions. Most of these originate from the application layer.}

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of libraries/framework reusable
methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
should not be documented on the method signature.  Then we inspected
the code for each method 
(defined either in the Android  Aplication Framework or in third-party libraries) 
explicitly signaling a runtime exception.

Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We found 79 methods (both from libraries and the Android platform) that  explicitly threw a runtime exception 
without listing it on the \emph{exception interface} (i.e., using 
throws clause in the method signature). From this set only one method (defined on a library)
included an @throws tag in its Javadoc - reporting that the given runtime exception
could be thrown in some conditions. These methods were responsible for 
267 exception stack traces mined in this study.

@@

\finding{%
Only a small fraction (4\%, 267 stack traces) of run time exceptions are programmatically
thrown. Almost none (0.4\%, just one) of these were documented.}

@@

\emph{\textbf{Missing Checked Exceptions on Exception Interfaces.}} 
Our exception stack trace analysis revealed an unexpected bug hazard: a 
checked exception thrown by a native method and not declared on the exception 
interface of these methods signaling them.

The native method in question was defined in
 the Android platform, which uses Java Native Invocation (JNI) to access 
native C/C++ code. This exception was thrown by the method getDeclaredMethods defined in java.lang.Class. 
The Java-side declaration of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.
 However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
violating the declaration. The Java compiler could not detect this violation, because it does 
not perform static exception checking on native methods.

This type of bug is hard to diagnose
because the developer usually does not have access to the native implementations. 
Consequently, since it is not expected by the programmer, when such method throws 
this exception, such an undocumented exception may remain
uncaught and cause the app to crash, or maybe mistakenly handled by subsumption.
The exception stack traces reporting this scenario actually correspond
to a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).
%Among the Android projects on which this exception stack trace was reported we can cite:
%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618
%stars).

\finding{%
For native methods, even checked exceptions can be
thrown without being documented on the exception interface.
}

\subsection{Exception Wrappings}

@@

@@

@@

We can also observe that some stack traces include successive cross-type wrappings, 
such as: Runtime - Checked - Runtime - Checked - Runtime - Checked -
Runtime. 
Hence, although some of these wrappings may be a result from design decisions, the mis-use of exception wrapppings may make the exception handling 
code more complex (e.g., the multiple wrappings) and error-prone,
 and lead to ``exception confusion''.

To illustrate this problem we can use one of the wrappings discussed above.
When the developer is confronted with a checked exception, the designer of the API is telling him/her 
to handle the exceptional condition (according to Java Specification and 
best practices). However, such exception may be wrapping an Error such
as an OutOfMemoryError, which indicates a resource deficiency 
that the program cannot possibly recover from). Hence, trying to
handle such an exception  
may lead the program to an unpredictable state.

\finding{%
Cross-type exception wrappings are common.
They violate the semantics of Java's original exception
design (e.g., when mapping unrecoverable Errors to other types of
exceptions),
and may lead to lengthy exception chains.
}

@@

%\noindent
\emph{Internal Validity.} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually mapped the concerns related
to exceptions.

 To ensure the quality of the analysis, we calculated the interrater agreement after three independent 
developers classified a randomly selected sample (of 25 exception 
types from the total of 100); the interater agreement was high (96\%).

\emph{External Validity.} Our work uses the GHTorrent dataset, which although 
comprehensive and extensive is not an exact replica of Github. 
However, the result of this study does not depend on the analysis of
a complete Github dataset. Instead, the goal of our study was to 
pinpoint \emph{bug hazards} on the exception-related code based on 
exception stack trace mining of a subset of projects.

We limited our analysis to a subset of existing open-source Android projects.
We are aware that the exception stack traces reported 
for commercial apps can be different from the ones found in this study, and that
this subset is a small percentage of existing apps.
Such threats are similar to the ones imposed to other empirical studies 
which also use free or open-source Android apps~\cite{Linar13,McDon13,Ruiz12}.
Moreover, several exception stack traces that support the findings of this study
refered to exceptions coming from methods defined on Android Application Framework
and third-party libraries.  Additionaly,  the \emph{bug hazards} observed in this study are due to
characteristics of Java exception model, which can impose challenges 
the robustness of not only to Android apps but also to other systems
 based on the same exception model.

@@

In this section, we present work that is related to the present paper, divided into
four categories: i) papers that use the information available on stack traces;
ii) empirical studies on the usage of Java Exceptions and its fault proneness;
and iii) tools to extract stack trace information from natural language artifacts
(e.g., issues and emails) and iv) empirical studies involving Android apps.

@@

The goal of this paper is to investigate 
to what extent stack trace information can reveal \emph{bug hazards} 
related to exception handling code that may lead to a decrease in
application robustness. 
To that end, we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code. Overall it included 6,005
exception stack traces.

Our first key contribution is
a novel approach and toolset (ExceptionMiner) for analyzing Java
  exception stack traces as occurring on GitHub and Google Code
  issues.
Our second contribution is
an empirical study of over 6000 actual stack traces,
  demonstrating that (1) half of the system crashes are due to errors
  in programming logic, with null pointer exceptions being most
  prominent;
  (2) Documentation for explicitly thrown RuntimeExceptions is almost
  never provided; 
  (3) Extensive use of wrapping leads to hard to understand chains
  violating Java's exception handling principles.

Our results shed light on common problems and bug hazards in Java
exception handling code, and call for tool support to help developers
understand their own and third party exception handling and wrapping logic.

>>>>

Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
on its repository. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
on the issues including exception stack traces. Regardless of the issue labels, every exception stack
trace may contain relevant information concerning the exception structure of the
projects analyzed, and therefore can reveal \emph{bug hazards} on the exception-related code.
Because of this, we opted for not restricting the analysis to just defect issues.

@@

\subsection{Android Apps in Google Code}
Google Code contains widely used open-source Android apps (e.g. K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on the Google Code issue tracker: \url{https://github.com/k9mail/k-9/wiki/LoggingErrors}.}).
However, differently from GitHub, Google Code does not provide an API to access the information related
 to hosted projects.\footnote{Google code used to provide a Web service to its repositories, but this was deactivated in June 2013 in what Google called a ``clean-up action''.}
To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigates 
through the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.
To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
(on the Google Code search interface) for the term ``android''. On January 2014, when we queried Google Code, this resulted in a list of 788  projects. This list comprised the seeds sent to our Crawler.

The Crawler retrieved all issues and comments for these projects.
From this set, 724 projects defined at least 1 issue. Running the ExceptionMiner tool 
 in this set we observed that 183 projects had at least one issue containing an exception stack trace.
 Then we performed further clean up (similar to the one described previously) inspecting the site 
of each project. As a result we could identify 157 Android projects.  This set contained  127,456 issues in total,
 from which 1,963 exception stack traces were extracted. Table~\ref{txab:lables} illustrates the occurrences of different labels 
on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
the issues were labeled as ``Defect''. However, based on the same assumption described for the GitHub repository
 we considered all issues reporting stack traces (regardless its labels).

@@

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace needs to be classified according to its type
(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of the exceptions reported 
on stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
a base exception type. Hence in this study the bytecode analysis was used to discover the type 
of each mined exception when the jar file of such exception was available on the project or 
on a reused  Java library. An specific implementation (based on source code analysis)
 was needed to discover the exception type when the bytecode was not available.
With this module we analyzed all exceptions defined in the Android platform (Version 4.4, API level 19),
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover, we also analyzed the 
exceptions reported on stack traces that were defined on applications and third-party libraries 
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e., Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristics adopted in this classification.
To conduct this classification, we provide this module with the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application. To discover the packages for the first two origins
we can use to the Android specification.
To discover the packages for the third origin, the application itself, this module
extracts the manifest files of each Android app, which defines the main packages that the applications consist of.
 If this file is not available, the tool recursively analyzes the 
structure of source-code directories composing the appm filtering out the cases in which the application also includes the source code of reused libraries.
Then, based on this information and using pattern matching 
between the signaler name and the packages, this module identifies 
the origin of the exception signalers.

@@

 The NullPointerException was mainly signaled inside the application code (50\%) and the Android platform (31.5\%),
 although we could also find the NullPointerException being signaled by third-party libraries (16.3\%). 
Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or a bad practice to 
explicitly throw a NullPointerException. Some prefer to encapsulate such an exception on
an instance of IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
can be signalled by an API expecting a non-null argument.The high prevalence of NullPointerException is aligned with the
findings of other 
research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher,kechagia2014}. 
For instance, Kechagia and Spinellis showed that the NullPointerException was the  
most reported exception on the crash reports sent to BugSense (a bug report 
management service for Android applications)  \cite{kechagia2014}.
Other research on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} shows that most of the automatically 
detected bugs were due to NullPointerException and exceptions
implicitly-signaled by the Java 
environment due to programming mistakes or resource limitations
 (as the ones found in our study).

%Sunghun et
%al.~\cite{kim2013predicting} show that 38\% the bugs related to
%exception handling in the Eclipse project
%are caused by NullPointerException.
%Furhtermore,

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed a manual inspection in order to identify the underlying
concerns related to the most frequently reported root exceptions.
Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery)~\cite{Zhang12}. Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
the concern identification of each exception was based on intended
meaning of the particular exception type, as defined in 
its Javadoc documentaion and in the Java specification. 
For example: (i) an instance of ArrayOutOfBoundException  
refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all 
exceptions related to backward compatibility~\cite{javaback}, such as
InstantiationError, VerifyError, and IllegalAccessError.

@@

This analysis revealed that approximately 75\% of the exceptions
 that caused the stack traces are implicitly thrown by runtime 
environment  due to mistakes on programming logic (e.g., out-of-bounds array index, null pointer 
references) and resource limitations. Although such exceptions do not directly point to violations on
the best practices described before (which are related to the 
explicitly thrown exceptions) they impose a major threat to apps robustness,
and therefore represent a critical bug hazard to the exception 
handling code of Android applications. 
Security and concurrency, which are known to be critica issues on  Android apps, 
raised few of the reported exceptions (less then 5\% of the anlysed stack traces).

%\finding{%
%More than 50\% of the uncaught exceptions analyzed in this study are due to errors in
%programming logic, with the NullPointerException as most prevalent
%exception.
%For another 25\% the root cause relates to resource constraints.
%}

\finding{%
Exceptions related to programming logic and resource limitations represent a major bug hazard to Android apps - 
since they represent approximately 75\% of the exceptions that caused the stack traces.
From this set the NullPointerException as most prevalent exception.
}

\subsection{Can the exception types reveal bug hazards?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

We can observe that most of the reported exceptions are of type runtime 
(64.85\%); and that the most common origins are methods defined either on the Application (47.3\%)
or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
 We can also see, from Table~\ref{tab:typeroottab}, that in contrast with the other origins, most of the 
 exceptions signaled on Android Libcore (i.e., the set of libraries reused by Android) are 
checked exceptions. This set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
org.apache.http, org.json, org.xml, and javax. Signaling checked exceptions is considered a 
good practice (see best practice IV in Section~\ref{sec:best}) because by using 
checked exceptions a library can define a precise 
\emph{exception interface}~\cite{miller1997issues} to its clients.
 Since such libraries are widely used  in several projects, this
 finding can be attributed to the libraries' maturity.

@@

\finding{Almost two thirds of all crashes come from runtime
  exceptions. Most of these originate from the application layer.
Hence, most of the bug hazards are located on this layer.}

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of libraries/framework reusable
methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
should not be documented on the method signature.  Then we inspected
the code for each method 
(defined either in the Android  Aplication Framework or in third-party libraries) 
explicitly signaling a runtime exception. 
Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We found 79 methods (both from libraries and the Android platform) that  explicitly threw a runtime exception 
without listing it on the \emph{exception interface} (i.e., using 
throws clause in the method signature). From this set only one method (defined on a library)
included an @throws tag in its Javadoc - reporting that the given runtime exception
could be thrown in some conditions. These methods were responsible for 
267 exception stack traces mined in this study.

@@

\finding{%
Only a small fraction (4\%, 267 stack traces) of runtime exceptions are programmatically
thrown. Almost none (0.4\%, just one) of these were documented. Such undocumented 
runtime exceptions violate best practices III and IV reveal a bug harzard to Java/Android apps.}

@@

\emph{\textbf{Missing Checked Exceptions on Exception Interfaces.}} 
Our exception stack trace analysis revealed an unexpected bug hazard: a 
checked exception thrown by a native method and not declared on the exception 
interface of these methods signaling them. The native method in question was defined in
 the Android platform, which uses Java Native Invocation (JNI) to access 
native C/C++ code. This exception was thrown by the method getDeclaredMethods defined in java.lang.Class. 
The Java-side declaration of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.
 However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
violating the declaration. The Java compiler could not detect this violation, because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usually does not have access to the native implementations. 
Consequently, since it is not expected by the programmer, when such method throws 
this exception, such an undocumented exception may remain
uncaught and cause the app to crash, or maybe mistakenly handled by subsumption.
The exception stack traces reporting this scenario actually correspond
to a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).
%Among the Android projects on which this exception stack trace was reported we can cite:
%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618
%stars).

\finding{%
For native methods, even checked exceptions can be
thrown without being documented on the exception interface.
It violates best practices III and IV and represent a bug hazard
 hard to diagnose.
}

\subsection{Can the exception wrappings reveal bug hazards?}

@@

%{\footnotesize
%\begin{verbatim}
%   try {
%      ...
%   } catch (InterruptedException e) {
%      android.util.Log.w(..., e);
%   } catch (ExecutionException e) {
%      throw new RuntimeException("...",e.getCause());
%   } catch (CancellationException e) {
%      ...
%   } catch (Throwable t) {
%     throw new RuntimeException("...", t);
%   }
%\end{verbatim}
%}

@@

\finding{%
A common practice in Android plaform are general catch clauses masking any exception 
into a general RuntimeException - it violates best practice III and is considered a bug 
hazard as it loses contextual information about the exception.
}

@@

We can also observe that some stack traces include successive cross-type wrappings, 
such as: Runtime - Checked - Runtime - Checked - Runtime - Checked -
Runtime. 
Hence, although some of these wrappings may be a result from design decisions, the mis-use of exception wrapppings may make the exception handling 
code more complex (e.g., the multiple wrappings) and error-prone,
 and lead to ``exception confusion''. To illustrate this problem we can use one of the wrappings discussed above.
When the developer is confronted with a checked exception, the designer of the API is telling him/her 
to handle the exceptional condition (according to Java Specification and 
best practices). However, such exception may be wrapping an Error such
as an OutOfMemoryError, which indicates a resource deficiency 
that the program cannot possibly recover from). Hence, trying to
handle such an exception  
may lead the program to an unpredictable state.

\finding{%
Cross-type exception wrappings are common.
They represent a bug hazard once they violate the semantics of Java's original exception
design, detailed on best practices I and II  (e.g., when mapping unrecoverable Errors to other types of
exceptions). 
}

@@

%\noindent
\emph{Internal Validity.} We used a heuristics-based parser to mine
exceptions from issues.  Our parsing strategy was conservative by default; for
example, we only considered exception names using a fully qualified class name
as valid exception identifiers, while, in many cases, developers use the
exception name in issue description. Conservative parsing may minimize false
positives, which was our initial target, but also tends to increase false
negatives, which means that some cases may have not been identified as
exceptions or stack traces. Our limited manual inspection did not reveal such
cases. Moreover, in this study we manually mapped the concerns related
to exceptions. To ensure the quality of the analysis, we calculated the
 interrater agreement after three independent 
developers classified a randomly selected sample (of 25 exception 
types from the total of 100); the interater agreement was high (96\%).

\emph{External Validity.} Our work uses the GHTorrent dataset, which although 
comprehensive and extensive is not an exact replica of Github. 
However, the result of this study does not depend on the analysis of
a complete Github dataset. Instead, the goal of our study was to 
pinpoint \emph{bug hazards} on the exception-related code based on 
exception stack trace mining of a subset of projects.
We limited our analysis to a subset of existing open-source Android projects.
We are aware that the exception stack traces reported 
for commercial apps can be different from the ones found in this study, and that
this subset is a small percentage of existing apps.
Such threats are similar to the ones imposed to other empirical studies 
which also use free or open-source Android apps~\cite{Linar13,McDon13,Ruiz12}.
Moreover, several exception stack traces that support the findings of this study
refered to exceptions coming from methods defined on Android Application Framework
and third-party libraries.  Additionaly,  the \emph{bug hazards} observed in this study are due to
characteristics of Java exception model, which can impose challenges 
the robustness of not only to Android apps but also to other systems
 based on the same exception model.

@@

In this section, we present work that is related to the present paper, divided into
four categories as detailed next.

@@

The goal of this paper is to investigate 
to what extent stack trace information can reveal \emph{bug hazards} 
related to exception handling code that may lead to a decrease in
application robustness. 
To that end, we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code. Overall it included 6,005
exception stack traces. Our first key contribution is
a novel approach and toolset (ExceptionMiner) for analyzing Java
  exception stack traces as occurring on GitHub and Google Code
  issues.
Our second contribution is
an empirical study of over 6000 actual stack traces,
  demonstrating that (1) half of the system crashes are due to errors
  in programming logic, with null pointer exceptions being most
  prominent;
  (2) Documentation for explicitly thrown RuntimeExceptions is almost
  never provided; 
  (3) Extensive use of wrapping leads to hard to understand chains
  violating Java's exception handling principles.
Our results shed light on common problems and bug hazards in Java
exception handling code, and call for tool support to help developers
understand their own and third party exception handling and wrapping logic.
